---
title: "DSCI 522 Milestone 1"
subtitle: Inferential Analysis for Egg Production
format:
  html:
    toc: true
editor_options:
  markdown:
    wrap: none
---

```{r load libraries, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(infer)
```

# Summary

Type Summary here.

# Introduction

Type Introduction here.

# Methods

## Data

Type here.

## Analysis

Type here.

# Result & Discussion

### Define the Question

Is there a difference in the average number of eggs produced by each hen between hatching and table eggs?

### Specify the Hypotheses and Observed Test Statistic

$H_{0}$: The mean number of eggs produced by each hen for hatching, $\mu_{hatching}$, is equal to the mean number of eggs produced by each hen for table eggs, $\mu_{table}$. Mathematically, $\mu_{hatching} = \mu_{table}$.

$H_{A}$: The mean number of eggs produced by each hen for hatching, $\mu_{hatching}$, is NOT equal to the mean number of eggs produced by each hen for table eggs, $\mu_{table}$. Mathematically, $\mu_{hatching} \ne \mu_{table}$.

*where* $\mu_{hatching}$ *represents the population mean number of eggs produced by each hen for hatching, and* $\mu_{table}$ *represents the population mean number of eggs produced by each hen for table eggs.*

The observed test statistic $\delta^{*}$ = $\hat{\mu}_\text{hatching} - \hat{\mu}_\text{table}$

*The estimate* $\delta^{*}$ *is the difference of the observed mean number of eggs produced by each hen for hatching and the observed mean number of eggs produced by each hen for table eggs coming from our egg production data set.*

### Hypothesis Test Using the Permutation Method

```{r Read Data}
# read in data
egg_prod <- read_csv("../data/egg-production.csv")
egg_prod
```

```{r Data Wrangling}
# create a new column to calculate the mean number of eggs produced by each hen
# Since cage-free is only available for table eggs, we filter the prod_process is equal to all 
egg_prod <- egg_prod |>
  filter(prod_process == "all") |>
  mutate(n_egg_by_hen = n_eggs / n_hens)
egg_prod
```

```{r Calculate the Test Statistic}
# manipulate the data to calculate the test statistic
n_eggs_by_hen_set <- egg_prod |>
  group_by(prod_type) |>
  summarise(n = n(),mean_n_eggs_by_hen = mean(n_egg_by_hen, na.rm = TRUE))
test_stat <- diff(n_eggs_by_hen_set$mean_n_eggs_by_hen)
n_eggs_by_hen_set
test_stat
```

```{r Hyhothesis Test, message=FALSE, warning=FALSE}
# set prod_type as factor
egg_prod <- egg_prod |>
  mutate( prod_type = factor(prod_type))

# implement Hypothesis test using the permutation
set.seed(522)
null_distribution <- egg_prod |>
  specify(formula = n_egg_by_hen ~ prod_type) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", order = c("table eggs", "hatching eggs"))

# calculate the p-value
p_value <- null_distribution |>
  get_pvalue(obs_stat = test_stat, direction = "both")

p_value
```

# References

Type here.
