---
title: "DSCI 522 Milestone 1"
subtitle: Exploratory Data Analysis for Egg Production
output:
 html_document:
  toc: yes
  toc_depth: '2'
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, message=FALSE, warning=FALSE}
library(infer)
library(digest)
library(ggplot2)
library(ottr)
library(tidyr)
library(dplyr)
library(forcats)
library(purrr)
library(readr)
library(knitr)
library(cowplot)
library(datateachr)
```

Import the data:
```{r}

library(readr)
egg_production <- read_csv("../data/egg-production.csv")

```
Step 1: Looking at the first and last few rows of the data. Finding the column data types and the NA values if any. 

```{r}
summary(egg_production)
str(egg_production)
head(egg_production)
tail(egg_production)
nrow(egg_production)

```

Comment: The source column should not be in the data frame for the analysis. n_hens and n_eggs are both numerical data types where as prod_type and prod_process are categorical data types. We can also take note that the n_hens has 9 NA values. The observed month gives us the time frame of when the data was recorded. We can also see that we have a relatively low number of rows: 320.


Step 2: Visualize the distribution of both n_hens and n_eggs columns as well as the correlation between them. 

```{r}
ggplot(egg_production,aes(x = n_eggs)) + geom_histogram() + labs(title = "Histogram of number of eggs in a month",
       x = "number of eggs in that month",
       y = "Count of particular bucket")

ggplot(egg_production,aes(x = n_hens)) + geom_histogram() + labs(title = "Histogram of number of hens",
       x = "number of hens in that month",
       y = "Count of particular bucket")

df <- na.omit(egg_production)

cor(df$n_hens, df$n_eggs)

```

Comment: We can see it is hard to determine a specific shape of the distributions but this is expected because we have a relatively small sample size. We also see that both distributions are somewhat similar. The correlation between the 2 variables is high at 0.9979883 meaning there is likely some strong relation between n_hens and n_eggs.


Step 3: Visualize the distribution of both prod_type and prod_process.

```{r}
ggplot(egg_production,aes(x = prod_type)) + geom_bar() + labs(title = "Histogram of production type",
       x = "different types of products",
       y = "Count")

ggplot(egg_production,aes(x = prod_process)) + geom_bar() + labs(title = "Histogram of production process",
       x = "different types of process",
       y = "Count")
``` 


Comment: In our data we have a higher production of tables eggs than hatching eggs. Regarding the processes we have much more all process than other kinds of process and about the same count for the cage-free organic vs non organic.



